Viral Sequence Sorting Tools - Rules on Fragments

### 2023-12-13 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updating SI2 and 5 to match figure in the main body of the text.

## Visualize how the precision, recall, and MCC changes across pipelines.
precision
```{r}
accuracy_scores <- accuracy_scores[order(accuracy_scores$precision, decreasing=F),]
accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
                                  color=ruletype, fill=ruletype)) +
  geom_boxplot(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Precision") +
  scale_fill_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
  scale_color_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6))
```

@@

recall
```{r}
accuracy_scores <- accuracy_scores[order(accuracy_scores$recall, decreasing=F),]
accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
                                  color=ruletype, fill=ruletype)) +
  geom_boxplot(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Recall") +
  scale_fill_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
  scale_color_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6))
```

@@

```
MCC
```{r}
mean_MCC <- accuracy_scores %>% 
  select(MCC, toolcombo) %>%
  group_by(toolcombo) %>%
  summarise(mean = mean(MCC)) %>%
  arrange(mean)
```

@@

```{r}
ggplot(confusion_vir_called, aes(y=mean, x=size,
                   fill=seqtype,
                   color=seqtype)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1)) +
  xlab("Length") +
  ylab("Sequences Called Viral (%)") 
```
```{r}
one.way <- aov(mean ~ seqtype, data = confusion_vir_called)

>>>>

## Visualize how the precision, recall, and MCC changes across pipelines.
precision
```{r}
mean_precision <- accuracy_scores %>% 
  select(precision, toolcombo) %>%
  group_by(toolcombo) %>%
  dplyr::summarise(mean = mean(precision)) %>%
  arrange(mean)
```

```{r}
accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = mean_precision$toolcombo)

#accuracy_scores <- accuracy_scores[order(accuracy_scores$precision, decreasing=F),]
#accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))

@@

fig <- ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
                                  color=ruletype, fill=ruletype)) +
  geom_boxplot(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Precision") +
  scale_fill_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
  scale_color_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6))

ggsave(
  "../IntermediaryFiles/all_precision_scatterplot.png",
  plot = fig,
  scale = 1,
  width = 8,
  height = 4,
  units = c("in", "cm", "mm", "px"),
  dpi = 300,
  limitsize = TRUE
)
```

```{r}
ann_col = data.frame(
    tnv = combos_list$tune_not_viral,
    tv = combos_list$tune_viral,
    dvf = combos_list$DVF,
    vb = combos_list$VIBRANT,
    vs = combos_list$VS,
    vs2 = combos_list$VS2
)
rownames(ann_col) <- combos_list$toolcombo

#viral_scores_1 <- accuracy_scores %>% filter(testing_set_index==5)
viral_scores_1 <- mean_precision
rn <- viral_scores_1$toolcombo

viral_scores_1 <- viral_scores_1 %>% select(c("mean"))
viral_scores_1 <- as.matrix(viral_scores_1)
#viral_scores_1 <- viral_scores_1 %>% select(c("mean"))
rownames(viral_scores_1) <- rn

pheatmap::pheatmap(viral_scores_1,
                  annotation_row = ann_col,
                  annotation_colors = ann_colors,
                  border_color = NA,
                  show_rownames = T,
                  show_colnames = F,
                  annotation_legend = F,
                  cluster_rows = F,
                  cluster_cols = F,
                  cellheight=3,
                  filename = "../IntermediaryFiles/heatmap_precision_labels_S2A.png"
                  )
```

@@

recall
```{r}
mean_recall <- accuracy_scores %>% 
  select(recall, toolcombo) %>%
  group_by(toolcombo) %>%
  dplyr::summarise(mean = mean(recall)) %>%
  arrange(mean)
```

```{r}
#accuracy_scores <- accuracy_scores[order(accuracy_scores$recall, decreasing=F),]
#accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))

accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = mean_recall$toolcombo)

@@

fig <- ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
                                  color=ruletype, fill=ruletype)) +
  geom_boxplot(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Recall") +
  scale_fill_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
  scale_color_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6))

ggsave(
  "../IntermediaryFiles/all_recall_scatterplot.png",
  plot = fig,
  scale = 1,
  width = 8,
  height = 4,
  units = c("in", "cm", "mm", "px"),
  dpi = 300,
  limitsize = TRUE
)
```

making heatmap for the x-axis label - Figure S2B
```{r}
ann_col = data.frame(
    tnv = combos_list$tune_not_viral,
    tv = combos_list$tune_viral,
    dvf = combos_list$DVF,
    vb = combos_list$VIBRANT,
    vs = combos_list$VS,
    vs2 = combos_list$VS2
)
rownames(ann_col) <- combos_list$toolcombo

#viral_scores_1 <- accuracy_scores %>% filter(testing_set_index==5)
viral_scores_1 <- mean_recall
rn <- viral_scores_1$toolcombo

#viral_scores_1 <- viral_scores_1 %>% select(c("recall"))
viral_scores_1 <- viral_scores_1 %>% select(c("mean"))
rownames(viral_scores_1) <- rn

pheatmap::pheatmap(as.matrix(viral_scores_1),
                  annotation_row = ann_col,
                  annotation_colors = ann_colors,
                  border_color = NA,
                  show_rownames = T,
                  show_colnames = F,
                  annotation_legend = F,
                  cluster_rows = F,
                  cluster_cols = F,
                  cellheight=3,
                  filename = "../IntermediaryFiles/heatmap_recall_labels_S2B.png"
                  )
```

@@

MCC
```{r}
mean_MCC <- accuracy_scores %>% 
  select(MCC, toolcombo) %>%
  group_by(toolcombo) %>%
  dplyr::summarise(mean = mean(MCC)) %>%
  arrange(mean)
```

@@

```{r}
ggplot(confusion_vir_called, aes(y=mean, x=size,
                   fill=seqtype,
                   color=seqtype)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1)) +
  xlab("Length") +
  ylab("Sequences Called Viral (%)") 
```

```{r}
one.way <- aov(mean ~ seqtype, data = confusion_vir_called)

### 2023-12-13 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd
Updating SI2 and 5 to match figure in the main body of the text.

ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 1))
```
```{r}
mean_MCC <- accuracy_scores %>% 
  select(MCC, toolcombo) %>%
  group_by(toolcombo) %>%
  summarise(mean = mean(MCC))
```

>>>>

ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 1))
```

making heatmap for the x-axis label - Figure S5
```{r}
ann_col = data.frame(
    tnv = combos_list$tune_not_viral,
    tv = combos_list$tune_viral,
    dvf = combos_list$DVF,
    vb = combos_list$VIBRANT,
    vs = combos_list$VS,
    vs2 = combos_list$VS2
)
rownames(ann_col) <- combos_list$toolcombo

ann_colors = list(
  tv = c("black", pal(6)[3]),
  tnv = c("black", pal(6)[1]),
  dvf = c("black", pal(6)[2]),
  vb = c("black", pal(6)[4]),
  vs = c("black", pal(6)[5]),
  vs2 = c("black", pal(6)[6])
)

viral_scores_1 <- accuracy_scores %>% filter(testing_set_index==5)
#viral_scores_1 <- mean_MCC
rn <- viral_scores_1$toolcombo

viral_scores_1 <- viral_scores_1 %>% select(c("MCC"))
#viral_scores_1 <- viral_scores_1 %>% select(c("mean"))
rownames(viral_scores_1) <- rn

pheatmap::pheatmap(as.matrix(viral_scores_1),
                  annotation_row = ann_col,
                  annotation_colors = ann_colors,
                  border_color = NA,
                  show_rownames = T,
                  show_colnames = F,
                  annotation_legend = F,
                  cluster_rows = F,
                  cluster_cols = F,
                  cellheight=3,
                  filename = "../IntermediaryFiles/heatmap_MCC_labels_S5.png"
                  )
```

```{r}
mean_MCC <- accuracy_scores %>% 
  select(MCC, toolcombo) %>%
  group_by(toolcombo) %>%
  summarise(mean = mean(MCC))
```
